---
title: "Cell composition"
output: html_notebook
---


```{r}
library(tidyverse)
library(dbscan)
library(speckle)
library(limma)
cols<-c("#E41A1C","#536D9E","#459E6E","#727E76","#B6656D","#F29022","#F781BF","#C598C9",
        "#AD6D5E","#7D8285","#4385BF","#204175","#1B9E77","#98D486","#D2C92D","#EFAC4C",
        "#F47494","#D8227D","#910241","#18ABB8","#6ECDDC","#B970A2","#A82676","#655D9B",
        "#8CA77B","#17542A","#94A908","#C8DD34","#8FC9BF","white")

files<-list.files("/media/Lawrenson_Lab_NAS/uthscsa/group_data/Xenium_labels/pembro",full.names = T)
files<-files[str_detect(files,"labels.gz")]
files<-files[str_detect(files,"UTP",negate = T)]

labels<-lapply(files,data.table::fread)
files<-str_remove_all(files,".+bro.|bnk.+")
names(labels)<-files
labels<-lapply(files,function(x) labels[[x]]%>%mutate(Section=x))%>%bind_rows()
```

# Coordinates
```{r}
coords<-data.table::fread("/media/Lawrenson_Lab_NAS/uthscsa/group_data/Xenium_labels/pembro/Banksy_domains.gz")
coords5k<-data.table::fread("/media/Lawrenson_Lab_NAS/uthscsa/group_data/Xenium_labels/pembro/Banksy_domains5k.gz")

temp<-coords%>%select(x_centroid,y_centroid)%>%dbscan(eps = 100)
coords$circle<-temp$cluster
#meta<-meta%>%mutate(circle=ifelse(circle==5&y_centroid>6000,500,circle))
coords<-coords%>%mutate(Core_ID=case_when(circle%in%c(40:42)~'1',
                                      circle%in%c(1,38,39)~'2',#1
                                      circle%in%c(2,6,37)~'3',#1
                                      circle%in%c(36,14,13)~'4',#1
                                      circle%in%c(12,8,15)~'5',#1
                                      circle%in%c(7,29,35)~'10',#1
                                      circle%in%c(5,34,30)~'9',#1
                                      circle%in%c(16:18)~'7',#1
                                      circle%in%c(9:11)~'6',#1
                                      circle%in%c(28,27,25)~'11',#1
                                      circle%in%c(26,24,19)~'12',#1
                                      circle%in%c(21:23)~'13',#1
                                      circle%in%c(4,3,44)~'14',#1
                                      circle%in%c(74,75,69)~'19',#2
                                      circle%in%c(70,67,66)~'18',#2
                                      circle%in%c(61,60,54)~'17',#2
                                      circle%in%c(50,52,53)~'16',#2
                                      circle%in%c(51,47,49)~'15',#2
                                      circle%in%c(73,68,71)~'20',#2
                                      circle%in%c(72,59,58)~'21',#2
                                      circle%in%c(55:57)~'22',#2
                                      circle%in%c(62:64)~'23',#2
                                      circle%in%c(65,48,76)~'24',#2
                                      circle%in%c(82:84)~'26',#3
                                      circle%in%c(81)~'25',#3
                                      circle%in%c(89,85)~'28',#3
                                      circle%in%c(92,91)~'29',#3
                                      circle%in%c(80,79)~'30',
                                      circle%in%c(77,78)~'31',
                                      TRUE~"discarded"))
names(cols)<-sort(unique(coords$Core_ID))
coords%>%ggplot(aes(x=-x_centroid,y=-y_centroid,col=Core_ID))+geom_point(size=.1)+
  geom_text(data=coords%>%group_by(circle)%>%summarise(x_centroid=median(x_centroid),
                                                     y_centroid=median(y_centroid)),aes(label=circle),col="black")+
  guides(col=guide_legend(override.aes = list(size=3)))+scale_color_manual(values=cols)

temp<-coords5k%>%select(x_centroid,y_centroid)%>%dbscan(eps = 100)
coords5k$circle<-temp$cluster
coords5k<-coords5k%>%mutate(circle=ifelse(circle==4&y_centroid>6000,400,circle))
coords5k<-coords5k%>%mutate(Core_ID=case_when(circle<4~'1',
                                      circle%in%c(4,8,9)~'2',#1
                                      circle%in%c(5,6,400)~'3',#2
                                      circle%in%c(13:15)~'4',#3
                                      circle%in%c(11,12,16)~'5',#3
                                      circle%in%c(23:25)~'9',#4
                                      circle%in%c(20:22)~'7',#4
                                      circle%in%c(17:19)~'6',#4
                                      circle%in%c(46,45)~'10',#5
                                      circle%in%c(54,53,38)~'11',#5
                                      circle%in%c(37,39,40)~'12',#5
                                      circle%in%c(34:36)~'13',#5
                                      circle%in%c(55:57)~'16',#5
                                      circle%in%c(58:60)~'17',#5
                                      circle%in%c(48:50)~'18',#5
                                      circle%in%c(42:44)~'19',#5
                                      circle%in%c(41,51,52)~'20',#5
                                      circle%in%c(73,66,63)~'14',#6
                                      circle%in%c(72,65,64)~'15',#6
                                      circle%in%c(67:69)~'23',#6
                                      circle%in%c(71,75,76)~'24',#6
                                      circle%in%c(85,84,82)~'21',#7
                                      circle%in%c(81,77,79)~'22',#7
                                      circle%in%c(94,93,91)~'26',#8
                                      circle%in%c(90)~'25',#8
                                      circle%in%c(95,96)~'28',#8
                                      circle%in%c(89,104)~'29',#8
                                      circle%in%c(102,100)~'30',#8
                                      circle%in%c(101,103)~'31',
                                      TRUE~"discarded"))

coords<-coords%>%mutate(panel="Protein")
coords<-coords5k%>%mutate(panel="5k")%>%bind_rows(coords)

#coords%>%filter(panel=="Protein")%>%select(-panel)%>%data.table::fwrite("/media/Lawrenson_Lab_NAS/uthscsa/group_data/Xenium_labels/pembro/Banksy_domains.gz")
#coords%>%filter(panel=="5k")%>%select(-panel)%>%data.table::fwrite("/media/Lawrenson_Lab_NAS/uthscsa/group_data/Xenium_labels/pembro/Banksy_domains5k.gz")
```

# Plot
```{r}
cols<-c("#D8227D","#536D9E","#459E6E","#EFAC4C","#F781BF","#C598C9",
        "#4385BF","#18ABB8","#910241","#E41A1C","#8FC9BF","white")
names(cols)<-unique(labels$percl_toplabel)

labels<-labels%>%mutate(cell_id=paste0(Section,"_",cell))
labels<-coords%>%select(cell_id,x_centroid,y_centroid,panel,circle,Core_ID,responder)%>%inner_join(labels)
labels<-labels%>%add_count(panel,circle)%>%filter(n>500)

plots<-lapply(unique(labels$Core_ID),function(i)
  labels%>%filter(Core_ID==i)%>%ggplot(aes(x=-x_centroid,y=-y_centroid,col=percl_toplabel))+
    geom_point(size=.1)+facet_grid(cols=vars(panel),scales="free",space="free")+
    scale_color_manual(values=cols)+guides(col=guide_legend(override.aes = list(size=3)))+
    ggtitle(paste0("Core_ID=",i))+theme(legend.position = "n"))

labels%>%count(percl_toplabel)%>%ggplot(aes(x=n,y=percl_toplabel,fill=percl_toplabel))+geom_col()+geom_text(aes(label=n))+scale_fill_manual(values=cols)+theme(legend.position = "n")
```

# Contrast
```{r}
temp<-read_tsv("/media/Lawrenson_Lab_NAS/uthscsa/group_data/Xenium_labels/pembro/PEmbro_TMA_deidentified.tsv")
labels<-labels%>%mutate(Core_ID=as.numeric(Core_ID))%>%left_join(temp,by = c("Core_ID"="Core ID"))

temp<-labels%>%filter(panel=="Protein")
temp<-temp%>%mutate(Gr=ifelse(responder==T,"responder","non_responder"))
run_speckle<-function(metadata){
  props<-getTransformedProps(metadata$percl_toplabel,
                           metadata$StudyID)
  ids<-metadata%>%distinct(Gr,StudyID)
  designM<-model.matrix(~0+Gr,data=ids)#but Cores are not independent 

  props<-props$TransformedProps[,ids$StudyID]
  fit<-lmFit(props,designM)
  cont<-makeContrasts(Grresponder-Grnon_responder,levels = designM)
  fit<-contrasts.fit(fit,cont)
  fit <- eBayes(fit, robust=TRUE)
  res<-topTable(fit,number = nrow(props),coef = 1)
return(res)}
proteinres<-run_speckle(temp)
temp<-labels%>%filter(panel=="5k")
temp<-temp%>%mutate(Gr=ifelse(responder==T,"responder","non_responder"))
res5k<-run_speckle(temp)

temp<-proteinres%>%rownames_to_column("ct")%>%mutate(proteinLFC=logFC)%>%select(ct,proteinLFC)
temp<-res5k%>%rownames_to_column("ct")%>%mutate(LFC5k=logFC)%>%select(ct,LFC5k)%>%inner_join(temp)
temp%>%ggplot(aes(x=proteinLFC,y=LFC5k,col=ct,label=ct))+geom_point(size=3)+geom_text_repel()+scale_color_manual(values=cols)+geom_vline(xintercept = 0)+geom_hline(yintercept = 0)
```

# Verify markers
## 5k
```{r}
files<-list.files("/media/Xenium_On_NAS/xenium_reanalysis",full.names = T)
files<-files[str_detect(files,"TMA")]
files<-files[str_detect(files,"Prime")]
files<-files[str_detect(files,"UTP",negate = T)]

xenium.list<-lapply(files, function(x) Read10X_h5(paste0(x,"/outs/cell_feature_matrix.h5")))
names(xenium.list)<-str_remove(files,".+sis.")
xenium.list<-lapply(xenium.list,function(x) x$`Gene Expression`)
xenium.list<-lapply(xenium.list,function(x) CreateSeuratObject(counts = x))
inte<-merge(xenium.list[[1]],y = xenium.list[names(xenium.list)[-1]],
            add.cell.ids =names(xenium.list))
rm(xenium.list)

inte<-subset(inte,cells = labels$cell_id)
temp<-labels%>%filter(cell_id%in%colnames(inte))
temp<-temp[order(match(temp$cell_id,colnames(inte))),]
inte<-AddMetaData(inte,metadata = temp%>%column_to_rownames("cell_id"),col.name = "percl_toplabel")

j<-c("KIT","TPSAB1","SIGLEC8","PTPRC","CD79A","BANK1","MS4A1",
     "CD27","CD38","BCL6","CXCR5","MME","JCHAIN","PRDM1","SDC1",
     "IGHA1","IGHG1","IGHG2","IGHG3","IGHG4","IGHM","NCAM1","FCGR3A",
     "CD3D","TRDC","TRGC1","TRGC2","CD4","IL4","GATA3","CCR4","CXCR4",
     "RORC","IL17A","STAT3","CCR6","FOXP3","IL2RA","CD8A","SELL",
     "CCR7","CD44","IL7R","CD69","CTLA4","LAG3","PDCD1","TIGIT","HAVCR2",
     "TOX","CD1C","HLA-DRA","HLA-DPA1","ITGAM","CD68","CXCL10","CCL2",
     "NOS2","ARG1","CD163","IL10","EPCAM","KRT10","KRT18","KRT8","WT1",
     "MUC5B","SOX9","SPDEF","LGR5",
     "PAX8","ESR1","FOXJ1","PAEP","PECAM1","PGR","IGF1","MMP11","FOXO1",
     "DCN","PDGFRA","THY1","COL1A1","ACTA2","TAGLN","MYL9","FAP","C7","GAS5")
DotPlot(inte,group.by = "percl_toplabel",features = j)+
scale_color_gradient2(low="#2166ac", mid = "#f7f7f7", high = "#b2182b")+
  theme(axis.text.x = element_text(angle=45,vjust = 1,hjust = 1))

```

## Protein panel
```{r}
files<-list.files("/media/Xenium_On_NAS/xenium_reanalysis",full.names = T)
files<-files[str_detect(files,"Pro")]
files<-files[str_detect(files,"UTP",negate=T)]
xenium.list<-lapply(files, function(x) Read10X_h5(paste0(x,"/outs/cell_feature_matrix.h5")))

prote.list<-lapply(xenium.list,function(x) x$`Protein Expression`)
names(xenium.list)<-str_remove(files,".+sis.")
names(prote.list)<-str_remove(files,".+sis.")

xenium.list<-lapply(xenium.list,function(x) x$`Gene Expression`)
xenium.list<-lapply(xenium.list,function(x) CreateSeuratObject(counts = x))
inte<-merge(xenium.list[[1]],y = xenium.list[names(xenium.list)[-1]],
            add.cell.ids =names(xenium.list))
inte<-subset(inte,cells = labels$cell_id)
temp<-labels%>%filter(cell_id%in%colnames(inte))
temp<-temp[order(match(temp$cell_id,colnames(inte))),]
inte<-AddMetaData(inte,metadata = temp%>%column_to_rownames("cell_id"),col.name = "percl_toplabel")
DotPlot(inte,group.by = "percl_toplabel",features = j)+
    scale_color_gradient2(low="#2166ac", mid = "#f7f7f7", high = "#b2182b")+
    theme(axis.text.x = element_text(angle=45,vjust = 1,hjust = 1))

prote.list<-lapply(prote.list,function(x) CreateSeuratObject(counts = x))
inteP<-merge(prote.list[[1]],y = prote.list[names(prote.list)[-1]],
            add.cell.ids =names(prote.list))
inteP<-subset(inteP,cells = labels$cell_id)
inteP<-AddMetaData(inteP,metadata = temp%>%column_to_rownames("cell_id"),col.name = "percl_toplabel")
DotPlot(inteP,group.by = "percl_toplabel",features = rownames(inteP))+
    scale_color_gradient2(low="#2166ac", mid = "#f7f7f7", high = "#b2182b")+
    theme(axis.text.x = element_text(angle=45,vjust = 1,hjust = 1))
